---
title: "BatchQC Report"
author: "Solaiappan Manimaran"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    toc: true
    toc_depth: 2
    template: batchQC.html
    self_contained: no
    lib_dir: libs
---

```{r, echo=FALSE}
shinyInput <- getShinyInput()
dat <- shinyInput$data
lcounts <- log2CPM(dat)$y
report_option_vector <- shinyInput$report_option_vector
eval_cell_1 = (report_option_vector[1]=="1")
```
`r if (eval_cell_1) 'Data Summary'`
`r if (eval_cell_1) '============'`
`r if (eval_cell_1) 'This is a summary of the given data matrix for the first 12 genes.'`
```{r eval=eval_cell_1, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
datasummary <- summary(t(dat))
panderOptions("table.split.table", 120)  ## table split at 100 (default 80) characters in a line
pander(datasummary[,1:12])
```

```{r, echo=FALSE}
eval_cell_1_1 = (report_option_vector[1]=="1")
```
`r if (eval_cell_1_1) 'Sample Boxplots'`
`r if (eval_cell_1_1) '======='`
`r if (eval_cell_1_1) 'Boxplots for all values for each of the samples and are colored by batch membership.'`

```{r eval=eval_cell_1_1, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
boxplot(lcounts,col=as.numeric(as.factor(batch)),main="Sample Boxplots (colored by batch)",xlab="Samples")
```

```{r, echo=FALSE}
eval_cell_2 = (report_option_vector[2]=="1")
```
`r if (eval_cell_2) 'Heatmap'`
`r if (eval_cell_2) '======='`
`r if (eval_cell_2) 'This is a heatmap of the given data matrix showing the batch effects and variations with different conditions.'`
```{r, eval=eval_cell_2, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
batchqc_heatmap(dat, batch=batch, mod=mod)
```

```{r, echo=FALSE}
eval_cell_3 = (report_option_vector[3]=="1")
```
`r if (eval_cell_3) 'Sample Correlations'`
`r if (eval_cell_3) '==================='`
`r if (eval_cell_3) 'This is a heatmap of the correlation between samples.'`
```{r, eval=eval_cell_3, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
batchqc_correlation(dat, batch=batch, mod=mod)
```

```{r, echo=FALSE}
eval_cell_4 = (report_option_vector[4]=="1")
```
`r if (eval_cell_4) 'This plot helps identify outlying samples.'`
```{r, eval=eval_cell_4, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
batchqc_corscatter(dat, batch=batch, mod=mod)
```

```{r, echo=FALSE}
eval_cell_5 = (report_option_vector[5]=="1")
```
`r if (eval_cell_5) 'PCA: Principal Component Analysis'`
`r if (eval_cell_5) '================================='`
`r if (eval_cell_5) 'This is a plot of the top two principal components and showing the variation with respect to batch effects and different conditions.'`
```{r, eval=eval_cell_5, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
pca <- batchqc_pca(lcounts, batch=batch, mod=mod)
shinyInput <- getShinyInput()
shinyInput <- c(shinyInput, list("pc"=data.frame(pca$x), "vars"=pca$sdev^2))
setShinyInput(shinyInput)
```

```{r, echo=FALSE}
eval_cell_6 = (report_option_vector[6]=="1")
```
`r if (eval_cell_6) 'PCA Proportion Variation and correlation'`
`r if (eval_cell_6) '========================================'`
```{r, eval=eval_cell_6, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
res <- batchqc_pca_svd(lcounts, batch=batch, mod=mod)
```

```{r, echo=FALSE}
eval_cell_7 = (report_option_vector[7]=="1")
```
`r if (eval_cell_7) 'PCA Proportion Variation and correlation Table'`
`r if (eval_cell_7) '=============================================='`
```{r, eval=eval_cell_7, echo=FALSE, message=FALSE}
panderOptions("table.split.table", Inf)  ## don't split tables
pander(res)
```

```{r, echo=FALSE}
eval_cell_8 = (report_option_vector[8]=="1")
```
`r if (eval_cell_8) 'Combat Plots'`
`r if (eval_cell_8) '============'`
`r if (eval_cell_8) 'This is a plot showing whether parametric or non-parameteric prior is appropriate for this data. It also shows the Kolmogorov-Smirnov test comparing the parametric and non-parameteric prior distribution.'`
```{r, eval=eval_cell_8, echo=FALSE, message=FALSE}
kstest <- combatPlot(lcounts, batch=batch, mod=mod)
print(kstest)
```

```{r, echo=FALSE}
eval_cell_9 = (report_option_vector[9]=="1")
```
`r if (eval_cell_9) 'Batch Effects testing'`
`r if (eval_cell_9) '====================='`
`r if (eval_cell_9) 'This is a summary of the statistical test for batch effects.'`
```{r, eval=eval_cell_9, echo=FALSE, message=FALSE}
retval <- batchtest(pca, batch=batch, mod=mod)
```
