---
title: "batchQC Report"
author: "Mani"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
---
Heatmap
=======

This is a heatmap of the given data matrix showing the batch effects and variations with different conditions. 

```{r, echo=FALSE, message=FALSE}
batchqc_heatmap(data.matrix, batch=batch, mod=modmatrix)

```

PCA: Principal Component Analysis
=================================

This is a plot of the top two principal components and showing the variation with respect to batch effects and different conditions. 

```{r, echo=FALSE, message=FALSE}
pca <- batchqc_pca(data.matrix, batch=batch, mod=modmatrix)

```

PCA Proportion Variation and correlation
========================================

```{r, echo=FALSE, message=FALSE}
res <- batchqc_pca_svd(data.matrix, batch=batch, mod=modmatrix)
```

Batch Effects testing
=====================

This is a summary of the statistical test for batch effects. 

```{r, echo=FALSE, message=FALSE}
batchtest(pca, batch=batch, mod=mod)

```
