---
title: "batchQC Report"
author: "Solaiappan Manimaran"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    toc: true
    toc_depth: 2
---
Data Summary
============

This is a summary of the given data matrix for the first 12 genes. 

```{r, echo=FALSE, message=FALSE}
#{r, echo=FALSE, message=FALSE, results='asis'}
#knitr::kable(head(data.matrix[1:10]), format="html")
datasummary <- summary(data.matrix)
pander(datasummary[,1:12])
```

Heatmap
=======

This is a heatmap of the given data matrix showing the batch effects and variations with different conditions. 

```{r, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
batchqc_heatmap(data.matrix, batch=batch, mod=modmatrix)
```

PCA: Principal Component Analysis
=================================

This is a plot of the top two principal components and showing the variation with respect to batch effects and different conditions. 

```{r, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
pca <- batchqc_pca(data.matrix, batch=batch, mod=modmatrix)

```

PCA Proportion Variation and correlation
========================================

```{r, echo=FALSE, message=FALSE, fig.width = 8, fig.height = 8}
res <- batchqc_pca_svd(data.matrix, batch=batch, mod=modmatrix)
```

PCA Proportion Variation and correlation Table
==============================================

```{r, echo=FALSE, message=FALSE}
panderOptions("table.split.table", Inf)  ## don't split tables
pander(res)
```

Combat Plots
============
This is a plot showing whether parametric or non-parameteric prior is appropriate for this data. It also shows the Kolmogorov-Smirnov test comparing the parametric and non-parameteric prior distribution.

```{r, echo=FALSE, message=FALSE}
kstest <- combatPlot(data.matrix, batch=batch, mod=modmatrix)
print(kstest)
```

Batch Effects testing
=====================

This is a summary of the statistical test for batch effects. 

```{r, echo=FALSE, message=FALSE}
retval <- batchtest(pca, batch=batch, mod=mod)
```

